A higher order approximate static condensation method for multi-material diffusion problems

### 2018-11-28 num_sec.tex
Fix a typo in the sec name

\section{Numerical Results}\label{sec:num}

@@

The first set of tests is designed to check the property of the method to accurately recover the solution that is the polynomial of degree~1. Let us consider the diffusion problem~\eqref{dual} with~$c \equiv f \equiv 0$, $\partial\Omega_D = \partial\Omega$. The computational domain $\Omega=(0,1)^2$ is divided into several subdomains by non-intersecting straight lines. In these settings the interface reconstruction algorithm MOF reconstructs the interfaces exactly. To test the {\it linearity preservation} property, we set up a pseudo-multi-material problem with the diffusion tensor being the same in all subdomains, $\vect K_i = \vect I$. The exact solution a linear function. The geometry of two cases under consideration is shown in Figure~\ref{fig:fake}.

@@

\subsection{Algebraic robustness}
%
%	One may expect that the conditioning of the ASC($n$) system~\eqref{asc_system} gets worse as one increases~$n$. In fact, we have the following result.\\
%	\begin{theorem}\label{thm:eig}
%		Let~$\mu_{\text{min}}^n > 0$ be the minimal eigenvalue of the system matrix~$\vect S_\bmesh$ in~\eqref{asc_system} of the ASC($n$) method. Then we have
%		\begin{equation}\label{mineig}
%			\mu_{\text{min}}^1 < \mu_{\text{min}}^0.
%		\end{equation}
%	\end{theorem}
%	\begin{proof}
%		\Sasha{We have it. To be added.}
%	\end{proof}
%
%	\Sasha{We also believe that the max eig value stays the same (based on num experiments), i.e. $\mu_{\text{max}}^1 \approx \mu_{\text{max}}^0$ or even $\mu_{\text{max}}^1 = \mu_{\text{max}}^0$}. I think it would be nice to prove it here, so we can show that the cond number is governed by the smallest eig value alone.
%	
In this section, we study the dependence of the condition number of matrix $\vect S_\bmesh$ on the position of the material interface against the background mesh. For this purpose, we solve the diffusion problem~\eqref{dual} in the unit square with $\partial\Omega = \partial\Omega_D$, $\vect K = k\,\vect I$, $k = 1$ on the left part and $k = 0.1$ on the right. We keep the mesh fixed, and change the position of the interface so that the minimal length of mini-faces~$w$ gets smaller, $w = 10^{-1}, 10^{-2}, \dots, 10^{-5}$ (see Figure~\ref{fig:w}).

\begin{figure}[h]
	\centering
	\caption{Distribution of materials leads to  different values of the minimal length of mini-faces $w$ \label{fig:w}}
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew1.png}
		\caption{$w = .1$}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew01.png}
		\caption{$w = .01$}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew001.png}
		\caption{$w = .001$}
	\end{subfigure}
\end{figure}

@@

Numerical results are shown in Table~\ref{fig:w:res}. We observe that the condition number~$\kappa_{\text{ASC(0)}}$ of $\vect S_\bmesh$ for ASC(0) levels off if $w$ gets smaller. The condition number~$\kappa_{\text{ASC(1)}}$ of ASC(1) depends on $w$ and behaves as $O(w^{-1})$ for $w\to0$. A closer look at the spectrum of~$\vect S_\bmesh$ reveals that the growth of the condition number for ASC(1) is due to presence of only few (three for this example) small eigenvalues, which tend to zero. To illustrate this, Table~\ref{fig:w:res} shows the ``effective'' condition number of $\vect S_\bmesh$ that is defined as
$$
	\widetilde{\vect\kappa}_{\text{ASC(1)}} = {\mu_{\text{max}}}/{\mu_3}
$$
where $\mu_{\text{min}} = \mu_0 \ge \mu_1 \ge \dots \ge \mu_{\text{max}}$ are the eigenvalues of~$\vect S_\bmesh$ of ASC(1). From this results we see that the effective condition number of ASC(1) stays bounded with respect to the interface position and is  close to the condition number of ASC(0). We hypothesize that the number of outliers in the  spectrum of $\vect S_\bmesh$ is proportional to the number of multi-material cells with small cuts.

@@

We finally apply the ASC methods to simulate the time-dependent diffusion problem. In the mixed form, the problem reads
\begin{equation}\label{dual_transient}
	\arraycolsep=2pt\def\arraystretch{1.7}
	\left\{\begin{array}{rcrccl}
		\vect K^{-1}\,\vect u & + & \nabla\,p & = & 0 &\quad\text{in } \Omega , \\
		\nabla\cdot\vect u    & + & \frac{\partial}{\partial t}\,p      & = & f &\quad\text{in } \Omega
	\end{array}\right.
\end{equation}%
%\begin{equation}\label{dual_transient}
%		\left\{\begin{split}
%			\vect K^{-1}\,\vect u &+ \nabla\,p\,&= 0&\quad\text{in } \Omega , \\
%			\nabla\cdot\vect u    &+ \frac{\partial}{\partial t}\,p       &= f&\quad\text{in } \Omega,
%		\end{split}\right.
%\end{equation}
for $t\in(0, T]$ with initial data $p(\vect x, 0) = p_0(\vect x)$ and boundary data as in \eqref{dual:bc:neumann}. After discretizing in time by the implicit Euler method, the problem takes the form \eqref{dual}--\eqref{dual:bc:neumann} with $c=|\Delta t|^{-1}$ and we apply the spacial discretization described in section~\ref{sec:ASC}. For the purpose of comparison, we apply arithmetic and harmonic homogenization methods followed by a MFD discretization.

@@

From the figures it is easy to appreciate  that among all the methods only  ASC(1)  provides a reasonable approximation for the coarse mesh, cf. Figures~\ref{fig:transient:comp}.
Numerical solution computed with ASC(0) overestimates incoming fluxes for the ring. This solution converges to the equilibrium state~$p_h \approx 1$ at approximately $t = 3.75$, which is significantly earlier than the time when same state is achieved by the reference solution. Note that the coarse mesh contains faces with three materials, and we saw already that ASC(0) may fail to converge for this case even for stationary examples.

>>>>

\section{Numerical results}\label{sec:num}

@@

The first set of tests is designed to check the property of the method to accurately recover the solution that is the polynomial of degree~1. Let us consider the diffusion problem~\eqref{dual} with~$c \equiv f \equiv 0$, $\partial\Omega_D = \partial\Omega$. The computational domain $\Omega=(0,1)^2$ is divided into several subdomains by non-intersecting straight lines. In these settings the interface reconstruction algorithm MOF reconstructs the interfaces exactly. To test the {\it linearity preservation} property, we set up a pseudo-multi-material problem with the diffusion tensor being the same in all subdomains, $\vect K_i = \vect I$. The exact solution is a linear function. The geometry of two cases under consideration is shown in Figure~\ref{fig:fake}.

@@

\subsection{Algebraic robustness}
%
%	One may expect that the conditioning of the ASC($n$) system~\eqref{asc_system} gets worse as one increases~$n$. In fact, we have the following result.\\
%	\begin{theorem}\label{thm:eig}
%		Let~$\mu_{\text{min}}^n > 0$ be the minimal eigenvalue of the system matrix~$\vect S_\bmesh$ in~\eqref{asc_system} of the ASC($n$) method. Then we have
%		\begin{equation}\label{mineig}
%			\mu_{\text{min}}^1 < \mu_{\text{min}}^0.
%		\end{equation}
%	\end{theorem}
%	\begin{proof}
%		\Sasha{We have it. To be added.}
%	\end{proof}
%
%	\Sasha{We also believe that the max eig value stays the same (based on num experiments), i.e. $\mu_{\text{max}}^1 \approx \mu_{\text{max}}^0$ or even $\mu_{\text{max}}^1 = \mu_{\text{max}}^0$}. I think it would be nice to prove it here, so we can show that the cond number is governed by the smallest eig value alone.
%	
In this section, we study the dependence of the condition number of matrix $\vect S_\bmesh$ on the position of the material interface against the background mesh. For this purpose, we solve the diffusion problem~\eqref{dual} in the unit square with $\partial\Omega = \partial\Omega_D$, $\vect K = k\,\vect I$, $k = 1$ on the left part and $k = 0.1$ on the right. We keep the mesh fixed, and change the position of the interface so that the minimal length~$w$ of mini-faces gets smaller, $w = 10^{-1}, 10^{-2}, \dots, 10^{-5}$ (see Figure~\ref{fig:w}).

\begin{figure}[h]
	\centering
	\caption{Distribution of materials leads to  different values of the minimal length~$w$ of mini-faces \label{fig:w}}
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew1.png}
		\caption{$w = .1$}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew01.png}
		\caption{$w = .01$}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{.33\linewidth}
		\centering
		\includegraphicsw{skew001.png}
		\caption{$w = .001$}
	\end{subfigure}
\end{figure}

@@

Numerical results are shown in Table~\ref{fig:w:res}. We observe that the condition number~$\kappa_{\text{ASC(0)}}$ of $\vect S_\bmesh$ for ASC(0) levels off if $w$ gets smaller. The condition number~$\kappa_{\text{ASC(1)}}$ of ASC(1) depends on $w$ and behaves as $O(w^{-1})$ for $w\to0$. A closer look at the spectrum of~$\vect S_\bmesh$ reveals that the growth of the condition number for ASC(1) is due to presence of only few (three for this example) small eigenvalues, which tend to zero. To illustrate this, Table~\ref{fig:w:res} shows the ``effective'' condition number of $\vect S_\bmesh$ that is defined as
$$
	\widetilde{\vect\kappa}_{\text{ASC(1)}} = {\mu_{\text{max}}}/{\mu_3}
$$
where $\mu_{\text{min}} = \mu_0 \le \mu_1 \le \dots \le \mu_{\text{max}}$ are the eigenvalues of~$\vect S_\bmesh$ of ASC(1). From this results we see that the effective condition number of ASC(1) stays bounded with respect to the interface position and is  close to the condition number of ASC(0). We hypothesize that the number of outliers in the  spectrum of $\vect S_\bmesh$ is proportional to the number of multi-material cells with small cuts.

@@

We finally apply the ASC methods to simulate the time-dependent diffusion problem. In the mixed form, the problem reads
\begin{equation}\label{dual_transient}
	\arraycolsep=2pt\def\arraystretch{1.7}
	\left\{\begin{array}{rcrccl}
		\vect K^{-1}\,\vect u & + & \nabla\,p & = & 0 &\quad\text{in } \Omega , \\
		\nabla\cdot\vect u    & + & \frac{\partial}{\partial t}\,p      & = & f &\quad\text{in } \Omega
	\end{array}\right.
\end{equation}%
%\begin{equation}\label{dual_transient}
%		\left\{\begin{split}
%			\vect K^{-1}\,\vect u &+ \nabla\,p\,&= 0&\quad\text{in } \Omega , \\
%			\nabla\cdot\vect u    &+ \frac{\partial}{\partial t}\,p       &= f&\quad\text{in } \Omega,
%		\end{split}\right.
%\end{equation}
for $t\in(0, T]$ with initial data $p(\vect x, 0) = p_0(\vect x)$ and boundary data as in \eqref{dual:bc}. After discretizing in time by the implicit Euler method, the problem takes the form \eqref{dual}--\eqref{dual:bc} with $c=|\Delta t|^{-1}$ and we apply the spacial discretization described in section~\ref{sec:ASC}. For the purpose of comparison, we apply arithmetic and harmonic homogenization methods followed by a MFD discretization.

@@

From the figures it is easy to appreciate  that among all the methods only  ASC(1)  provides a reasonable approximation for the coarse mesh, cf. Figures~\ref{fig:transient:comp}.
Numerical solution computed with ASC(0) overestimates incoming fluxes for the ring. This solution converges to the equilibrium state~$p_h \approx 1$ at approximately $t = 3.75$, which is significantly earlier than the time when the same state is achieved by the reference solution. Note that the coarse mesh contains faces with three materials, and we saw already that ASC(0) may fail to converge for this case even for stationary examples.
